### 2024-03-26 thesis/draft.tex
temp storage

\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{subcaption}

\pagenumbering{arabic}
\linespread{1.5}
\setlength{\parskip}{12pt}
\pagestyle{fancy}
\fancyhf{}

\lhead{Ningrui Xie}
\chead{Thesis First Draft}
\rhead{ID: 400323416}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\newcommand{\perday}{\ensuremath{\mathrm{/day}}}

\begin{document}

\title{Mathematical Approaches for Simulating Epidemic Progression: Addressing Limitations of the Linear Chain Trick in ODE Models}

\author{Submitted to\\ Dr. Jonathan Dushoff 
\\McMaster University\\Hamiltion, Ontario, Canada L8S 4K1}
\date {March 22, 2024}
\maketitle

\centerline{Reported by}
\centerline{\textbf{Ningrui Xie}}
\centerline{\textbf{400323416}}

\newpage
\section{Abstract}

\section{Introduction}
Ordinary differential equations (ODEs) are extensively utilized for mathematical modeling in the biological sciences, offering a robust framework to accurately describe the temporal dynamics of systems. Their capacity to model continuous state changes makes them essential in capturing complex biological phenomena such as population dynamics \cite{salisbury2011mathematical}\cite{sego2021generation}, epidemiological patterns of infectious diseases \cite{Anderson1991}\cite{Diekmann2000}\cite{Feng2016}, and cellular processes like proliferation in oncogenesis \cite{baker1998modelling}\cite{jarrett2018mathematical}. The inherent flexibility and simplicity of ODEs make them not only easy to solve and understand but also easy to manipulate and accommodate to various situations.

In the field of epidemiology, ordinary differential equations (ODEs) serve as a crucial tool for characterizing disease progression and making predictions. During the COVID-19 pandemic, they were extensively utilized to guide policy-making, especially in disease control and public health interventions \cite{thompson2020epidemiological}. A commonly used model is the SIR compartmental model \cite{Anderson1991}\cite{kermack1927contribution}, which divides the population into three states: Susceptible $(S)$, Infectious $(I)$, and Recovered $(R)$. In the model, a susceptible individual, upon contact with the infected population at a rate $\beta$ $(\frac{1}{time})$, transitions immediately to the infectious state. An infected individual then moves to the recoverd state at a recovery rate of $\gamma$ $(\frac{1}{time})$. However, one major limitation of this model it its implicit assumption that the duration of the infectious stage follows an exponential distribution. This can be observed in a simplified equation that models the dynamics of the infected population. It assumes the entire population was infected at the outset, and individuals can only transition out of the infectious stage at a rate of $\gamma$:
\begin{align}
    \frac{dI}{dt} = -\gamma I
\end{align}
Solving this equation explicitly yields
\begin{align}
    I(t) = I_0 e^{-\gamma t}
\end{align}
where $I_0$ represent the initial infected population. Formula (2) implies a stochastic state transition model where the time spend in the infectious $I$ is exponentially distributed with a mean stage duration of $\frac{1}{\gamma}$. This dwell time distribution suggests that an individual has the highest probability of recovering from the disease immediately after infection, which is often not realistic. In reality, a delay between the successful infection of an individual and the recovery from the disease is inevitable. The work conducted by Byrne AW et al. \cite{byrne2020inferred} presents the simulation of parameter distribution inferred for the duration of the infectious period, and the shape of this distribution closely resembles a gamma distribution. The study by Greenhalgh and Day (2017) \cite{greenhalgh2017time}, utilizing measles outbreak data, also demonstrates that the recovery rate from an infectious disease is not static but can vary based on a wide range of factors.

It is important to note that even when the mean is constant, the shape of the stage duration distributions (i.e., the timing of stage transitions) can significantly impact the dynamics and the outcomes of predictive models in applied settings \cite{krylova2013effects}\cite{keeling2002understanding}\cite{wearing2005appropriate}\cite{nguyen2008noise}. Consequently, it becomes crucial to consider the type of distribution characterizing an individual in a specific state when constructing these models. A standard approach for developing models that incorporate different stage duration distributions is through the use of integro-differential equations (IDEs) or integral equations (IEs) \cite{hurtado2019generalizations}\cite{kermack1927contribution}\cite{hethcote1980integral}. IDEs facilitate the extension of the SIR model to include arbitrary distributions for the duration of infectiousness, as discussed in various sources \cite{feng2000endemic}\cite{hethcote1980integral}\cite{ma2006generality}. This approach allows for the capture of the 'memory effect', where the current rate of change in each compartment of the model (Susceptible, Infectious, Recovered) is influenced by the entire history of the epidemic, rather than relying on a constant rate assumption.

However, while integro-differential equations increase the model's flexibility, they present challenges in formulation, mathematical analysis, and simulation, particularly in comparison to ODEs \cite{krylova2013effects}\cite{hurtado2019generalizations}\cite{burton2005volterra}. An alternative method, the Linear Chain Trick (LCT), overcomes some of these challenges, offering a simpler and more direct approach to modifying the shape of distributions \cite{macdonald1978time}\cite{smith2011introduction}. This approach subdivides the single infected state into n substages, each characterized by an identical exponential distribution with a rate of $n\gamma$. This technique is based on the principle that the sum of a series of independent exponentially distributed random variables follows an Erlang distribution \cite{krylova2013effects}\cite{therrien2018probability}, which is a special case of the gamma distribution with an integer shape parameter. By employing this method, the distribution of the infectious stage duration transitions from exponential to Erlang. The central tendency of the Erlang distribution towards its mean value offers the model a more accurate representation of real-world scenarios.

When applying LCT into the SIR model, the number of substages $n$ determines the shape of the stage duration distribution. As $n$ increases, the variance decreases, resulting in a distribution that is more peaked and narrow around the mean value $\frac{1}{\gamma}$. As $n$ approaches infinity, the distribution converges to a delta function (a deterministic value, implying that all individuals who become infectious at time $t$ recover at exactly time $t + \frac{1}{\gamma}$), and the system evolves into a delay differential equation \cite{krylova2013effects}\cite{hethcote1980integral}.

The Linear Chain Trick (LCT) is extensively employed in modeling biological processes, not only in quantifying disease progression \cite{lloyd2001destabilization}\cite{lloyd2001realistic}, but also in analyzing virus \cite{lloyd2001dependence}\cite{kakizoe2015method} and population \cite{cushing2013integrodifferential} dynamics. Studies have demonstrated that switching the stage duration distribution from an exponential to an Erlang distribution leads to predictions that align more closely with real data points \cite{kakizoe2015method}. However, despite the increased realism they bring to modeling disease processes, employing LCT in ODE models presents several limitations. One primary challenge is the inflexibility of the model's parameters. To integrate independent exponentially distributed random variables into the ODE system, $n$ must be an integer. As mentioned above, since the shape of the stage duration distribution is determined by $n$, it would also be constrained to discrete changes. Another inherent limitation of using LCT is the cumbersome and time-consuming process of determining the appropriate number of substage to divide. Multiple models with varying substages must be constructed and subjected to repeated data-fitting processes to identify the optimized parameter value. Prior research has investigated approaches for estimating the number of states in the linear chain, employing techniques such as profile likelihood \cite{raue2009structural} and model reduction \cite{maiwald2016driving}. Yet, a comparative study of these methods indicates that the typical quality of experimental data is insufficient for reliably estimating the chain length using these approaches \cite{hauber2020estimating}. These limitations significantly restrict the adaptability and practical application performance of the model.

The recently proposed Generalized Linear Chain Trick (GLCT) enhances the Linear Chain Trick (LCT) by enabling models to incorporate a broader phase-type family of distributions \cite{hurtado2019generalizations}\cite{hurtado2021building}\cite{bladt2017phase}, which includes exponential, Erlang, hypoexponential, and Coxian distributions, thus allowing for greater flexibility and the ability to capture more complexity. The GLCT achieves this by creating a higher-dimensional Markov chain, enabling a more accurate representation of the complex nature of dwell time distributions. However, this process can be mathematically heavy and challenging to interpret. Additionally, this approach may lead to a increase in the number of variables or states within the model, potentially escalating computational demands. Therefore, finding a simpler and more efficient method to enhance the flexibility of the ODE model remains essential.

In this study, we focused on one particular direction: generalizing the Erlang distribution of stage durations currently present in the model. The goals were to have a more effective way for parameter estimation and to allow for a more diverse range of shapes, achieved by eliminating the constraints imposed by the shape parameter ($\kappa$). To address these limitations, we proposed the Geometric Chain Trick (GCT). Instead of having a variable number of substages $(n)$ with a constant transition rate between them $(\gamma_i = n\gamma)$, we fix $n$ to a specific integer value and use a series of rates that follow a geometric pattern. This series is represented by the formula $\gamma_i = ar^{i-1}$, governed by parameters $a$ and $r$. This approach enables a uniform model structure with the parameter of interest transitioning from discrete values $(n$ and $ng)$ to continuous values $(a$ and $r)$, thereby enhancing the model's adaptability (details in Section 4.x).

\section{Method}

\section*{Results}

\section{Discussion}

%%%%%%%%

This is a citation \cite{Feng2016}

Another citation \cite{Anderson1991}

Here's a simple plot (with no dependence)

\begin{figure}[h!]
    \centering
    \includegraphics{figures/mytest1.Rout.pdf}
    \caption{Mytest1 plot (are you working??)}
\end{figure}

Another plot (depend on tempfunc.R)

\begin{figure}[h]
    \centering
    \includegraphics{figures/mytest2.Rout.pdf}
    \caption{Mytest2 plot}
\end{figure}

Maybe another plot from a different folder:

\begin{figure}[h]
    \centering
    \includegraphics{figures/test.Rout.pdf}
    \caption{test plot}
\end{figure}

\bibliographystyle{unsrt}
\bibliography{Reference}

\end{document}